
\section{Emerging  Applications}

\subsection{Entertainment}

In current times VoIP is reaching its capabilities as alternative to the traditional telephone service. Slowly the software industry is trying to integrate the VoIP protocol into their products. Currently integrated VoIP is used to enhance the user experience by allowing for intuitive and direct way of communicating with fellow users. And the biggest beneficent of this process is gaming industry. 

As high speed Internet connection became wide spread, the gaming industry began to focus on creating multiplayer games. Users playing early multiplayer games did not have a chance to communicate verbally; instead they were forced to write short messages through chat boxes. The introduction of VoIP allowed to overcome slow and cumbersome communication and created additional way of enhancing the game play\cite{website:voip_impact_on_gaming}.

However the early implementations of VoIP in multiplayer games were far from perfect. Jitter, lost packets, microphone echo effect, exiting from the game environment which resulted in the termination of voice chat were amongst main complaints from players. As a solution third party companies: Mumble, Ventrilo and Teamspeak, came up with more advanced implementations of VoIP for games. Allowing them to create their teams, adjust max input output volumes, regulate voice transmission thresholds and/or echo cancelation. But again there was a drawback to these solutions, to manage your voice communication players were forced to minimise the game prior to adjusting voice chat options.

Currently there is no golden solution to the above problems and gamers are forced to choose either functionality or ease of use. The gaming market changes rapidly though and there might be a possibility that VoIP services could be reliable, user-friendly and be integrated in all of games. There are two main gamimg platforms: consoles and the traditional PC.

First let's discuss consoles, as this is the fastest growing gaming sector\cite{website:shift_in_gaming_demographics}. Consoles by nature are a unified platform, can provide an integrated VoIP service and expose it to game developers so they have an option of utilising that console's VoIP implementation in their product. As the console market is very competitive all the main console companies: Sony, Microsoft and Nintendo are trying to develop good quality integrated voice services to gain a competitive advantage\cite{website:voip_adataion}.

The PC as a common platform is not so easy to consolidate as many different operating systems exist side-by-side. Building an integrated VoIP solution for all OS, old and new would be unfeasible. However in 2004 Valve Corporation released a platform called Steam. It provides an easy way to buy and store your games online. It also has an exposed API that provides a wrapper for any game it sales. Wrapper allows users to access Steam's resources such as friend's lists, groups or instant messaging. Steam grew popular over the last few years and became a predominant games distributor for the PC platform (estimated 70\% market share\cite{steam_market_share}). In 2007 Valve implemented a VoIP service. It's far from being perfect but the lack of a competitor does not stimulate Valve to further work on enhancing its implementation. Although Steam provides users with a clean interface, abilities to create "chat rooms" and manage transmission/receive volumes.

All of the platforms however were faced with one big problem how to serve a large client base without a large dedicated server farm and large bandwidth demand. In 2010 Steam reached 30 million users \cite{website:steam_growth}, Xbox Live 23 million users\cite{xbox_live_market} and Playstation3 33.5 million users\cite{website:ps3_marketshare}. To provide reliable VoIP services the platforms (similary to Skype) use a peer-to-peer architecture for providing VoIP. That way service providers are distributing workload to each client and only controlling how service is provided via authorisation and discovery services.

\subsection{Smart-phone Integration}

After establishing a solid foundation in the PC market VoIP took a natural step and began spreading to mobile devices. Data plans are now a common option for contract and pre-paid mobile users. Combined with a widespread 3G (3rd generation) mobile data network, handsets are equipped with fast data access that VoIP clients may take advantage of in order to allow mobile users gain cheaper calls.

However VoIP implementations are still suffering from many issues namely packet loss, delay and jitter. Despite that the biggest thing crippling VoIP is the handset manufacturers companies themselves. Apple's iPhone has restricted to VoIP transmission only on Wi-Fi networks, blocking 3G's data channel as means of VoIP communication, this of course limits where and when user can use VoIP. Microsoft with its Windows Mobile also limits the capabilities of VoIP, in contrary to iPhone it allows VoIP to access 3G but it limits access to internal handset earpiece. Resulting in bad quality of received voice, especially if external speaker is located on side or back of phone\cite{website:voip_cripling_by_apple}.

There is no definite statement by mobile producers why such a actions were taken. However one commonly shared opinion is that telecom companies which are closely cooperating with mobile producers to reduce the capabilities of VoIP in mobiles. To protect their market and force mobile customers to use only their services. This state of affairs would be maintined but the introduction of a new OS by Google is set to upset the market dynamics. Android is gaining popularity as it is an open-source product as it does not restrict the functionality of the device. As this alternative is gaining more and more mobile market share, proprietary OS's will have to remove their restrictions on VoIP to stay competitive.

Other issues that can really affect mobile VoIP is the lack of bandwidth. If the bandwidth pipe is too small for transmitting VoIP packets the audo quality of the conversation will be degraded. This is a real issue if VoIP is used over a Wi-Fi network which is also used by other users. Things are looking better in Vo3G (Voice over 3G), but there are still few concerns. Currently there are several concurrent applications running on a users' mobile phones and few of them require bandwidth to work. That is why 3G bandwidth can be an issue for a mobile VoIP user. The ideal bandwidth that 3G can reach is 2.4 Mb/s, however realistically a stationary user can expect to reach 2 Mb/s. That is enough for most of application running alongside VoIP application. But the faster user is moving the smaller bandwidth get. Pedestrian can expect 384Kb/s, vehicle passenger only 144Kb/s \cite{website:3g_spec}. 

Although most of VoIP codecs require less than 40 Kb/s and this is enough to support quality VoIP service\cite{website:voip_codecs}, but this shows that bandwidth can still be an issue if mobile user is running many application and happens to be in position where bandwidth is greatly reduced.


\subsection{VoIP and Video Conferencing}
Every business is looking to reduce the cost of operating and encourage its employees work more effectively. On top of that many companies have their branches spread across the whole world. VoIP has proven to be very successful as a cheap method of bringing people from across the world together. But this was not enough for businesses who demanded a new easier and more direct way for its employees to communicate. That is why high quality video conferencing is becoming quite popular across geographically diverse organisations. But there are issues and tradeoffs tied to this solution, ranging from technical to psychological. Usage of video conferencing software requires some ports (depending on the standard was used) to be open and since the connection to other site is based on p2p protocol, there is a possibility of a malicious actions taken by outside third party. H.323 protocol for example was not designed with security as its prime feature; many open ports increase a chance of attack on system\cite{website:vid_conf_voice_protocol_overview}.

Usually video conferencing is going to take place in offices however mobile video conferencing has already been implemented. As mentioned in section above VoIP alone needs as much as 40Kb/s of bandwidth, now we need to increase bandwidth requirements by the amount of bandwidth needed by video transmission. Naturally the bigger and better quality images are being sent/received the more bandwidth will be required. 
To psychological problems with video conferencing we can account the discomfort in delays of responses. Delays as small as 300ms can be spotted by users and create a feeling of discomfort and bad communication. Another issue is related to facial expressions, because of hardware limitations users are looking at the monitor rather than at the camera creating bad image in eye of the other person as it looks like person is avoiding an eye contact\cite{website:vid_conf_overview}.

\subsection{Unified Communications}

An underlying theme in the development of VoIP technologies is the integration with existing non-real-time forms of communication such as email, voicemail and fax services.

Initially focused at business customers unified communications can provide potentially great value when integrated with business processes, enabling decision makers to be more responsive to real-time data from production and work-flow systems no matter the persons location. The combining of presence information, messaging, conferencing, collaboration, email, calendaring and business systems with IP technology and the plethora of new form factor devices becomes a very powerful concept and provides new opportunities to communication system providers. A number of major firms are competing currently in this area namely Microsoft, Cisco, Google and IBM - with each firm attempting to merge it with their existing platforms.

The concept of unified communications is also making its way slowly into the domestic market, an early leader in this area was RIM with their Blackberry device which enabled mobile phone users to receive emails to their mobile device instantly. Recently with the increased adoption of IP on the mobile handset users are becoming familiar with sending and receiving Twitter messages, mobile email, uploading photos to social networks. While these communication streams have not yet be integrated they pave the way to having your presence and communications available on whatever device is appropriate at the time, with the network dynamically routing messages to the relevant location.

Recently Google has attempted to position itself as a unified communication provider with a new product called Google Voice. They offer integration of email, sms, traditional mobile phone services, VoIP, conferencing, switching calls to other devices mid conversation and numerous other features to users in both a desktop and mobile client.\cite{website:cellularNewsGoogleUC}

This area is still in its infancy, but with the increasing adoption of capable IP connected mobile handsets and the adoption cloud based services it appears to be a logical step forward.


